---
title: "Untitled"
format: html
editor: visual
execute: 
  warning: false
---



```{r}
library(tidyverse)
library(censored)
library(survival)

```


```{r}

bc <- ISLR2::BrainCancer |> 
  as_tibble()

```

Data:
- **status**: 0 alive, 1 dead  
- **gtv**: gross turmor volume (cubic cm)   
- **sex**: male or female  
- **diagnosis**: meningioma LG glioma, HG glioma, or other   
- **loc**: tumor location infratentorial or supratentorial  
- **ki**: Karnofsky index  
- **stereo**: stereotactic method; either stereotactic radiosurgery or fractionated stereotactic radiotherapy, abbreviated as SRS and SRT, respectively  


53 of 88 patients were still alive at the end. 

```{r}
table(bc$sex, bc$status)
```

Sex did not show statistical significance in difference with $p = 0.2$

```{r}

fit.surv <- survfit(Surv(time, status) ~ 1, data=bc)
plot(fit.surv, xlab='months', ylab='Est. prob of survival')

fit.sex <- survfit(Surv(time, status) ~ sex, data = bc)
plot(fit.sex, col = c(2, 4))
legend("bottomleft", levels(bc$sex), col = c(2,4), lty = 1)

logrank.test <- survdiff(Surv(time, status) ~ sex, data=bc)
logrank.test

fit.cox <- coxph(Surv(time, status) ~ sex, data = bc)
summary(fit.cox)

```

all predictors

```{r}
fit.all <- coxph(
  Surv(time, status) ~ 
    sex 
  + diagnosis
  + loc
  + ki 
  + gtv
  + stereo
  , data = bc
    )


fit.all


modaldata <- data.frame(
  diagnosis = levels(bc$diagnosis),
  sex = rep("Female", 4),
  loc = rep("Supratentorial", 4),
  ki = rep(mean(bc$ki), 4),
  gtv = rep(mean(bc$gtv), 4),
  stereo = rep("SRT", 4)
)


survplots <- survfit(fit.all, newdata = modaldata)

plot(survplots, col=2:5)
legend("bottomleft", levels(bc$diagnosis), col=2:5, lty = 1)

```

## Publication data

```{r}
pub <- ISLR2::Publication |> as_tibble()

fit.posres <- survfit(
  Surv(time, status) ~ posres, data = pub
)

fit.pub <- coxph(Surv(time, status) ~ posres, data = pub)
fit.pub

logrank.test.pub <- survdiff(Surv(time, status) ~ posres, data = pub)
logrank.test.pub


```

```{r}
fit.pub2 <- coxph(Surv(time, status) ~ . - mech, data=pub)
fit.pub2

fit.pub3 <- coxph(Surv(time, status) ~ posres  + clinend + budget + impact, data = pub)
fit.pub3
```
 
 
```{r}
ex_4 <- tibble(
  observation = c(26.5, 37.2, 57.3, 90.8, 20.2, 89.8),
  censor_ind = c(1, 1, 1, 0, 0, 0),
  covariate_x = c(0.1, 11, -0.3, 2.8, 1.8, 0.4)
)

coxph(Surv(observation, censor_ind) ~ covariate_x, data = ex_4) 

y <- survfit(Surv(observation, censor_ind) ~ 1, data = ex_4)
plot(y)


```

### Applied Exercises

$\delta$ 1 = true survival time, 0 = censored
$y$ time
$c$ censored time
$t$ survival time

 
```{r}

res <- survfit(Surv(time, status) ~ 1
               , data=bc)
plot(res)
```
 
```{r}

sample88 <- bc[, c('time', 'status')] |> sample_n(88, replace = TRUE)
sample200 <- bc[, c('time', 'status')] |> sample_n(200, replace = TRUE)

plot(res, col = 2)
lines(survfit(Surv(time, status) ~ 1, data = sample88), col = 3)
lines(survfit(Surv(time, status) ~ 1, data = sample200), col = 4)
legend("bottomleft", c('base', 'bootstrap_88', 'bootstrap_200'), col=2:5, lty = 1)

res88 <- survfit(Surv(time, status) ~ 1, data = sample88)
res200 <- survfit(Surv(time, status) ~ 1, data = sample200)

```
 
The Kaplan-Meier survival curve changes relative to the dataset:  

- Bootstrap $n=88$ leads to similar variance as before, only with slightly higher survival rates. This could be due to the random sampling receiving more censored data points or observations that survived longer.   
- Bootstrap $n=200$ had the similar survival line, with a much tighter distribution. This is a by-product of bootstrap sampling to increase the size of the data, but with the overall same characteristics.  
 
## Fit Cox PH

```{r}
fit.all <- coxph(
  Surv(time, status) ~ 
    sex 
  + diagnosis
  + loc
  + ki 
  + gtv
  + stereo
  , data = bc
    )

fit.all

```

The predictors HG glioma of diagnosis and ki being significant. Having a diagnosis of HG glioma corresponds to a multiplier of $exp(2.15) = 8.6$ in the instantaneous chance of dying. 

A one-unit increase in the Karnofsky index (ki) corresponds to a multiplier of $exp(-0.05) = 0.95$ in the instantaneous chance of dying. 
 
Predictors such as sex, loc, and stereo were not significant. 
 
 
```{r}
fit.all2 <- coxph(
  Surv(time, status) ~ diagnosis + ki + gtv, data = bc
)

fit.all2
```
 
```{r}
bc2 <- bc |> mutate(ki = ifelse(ki == 40, 60, ki))
modaldata2 <- data.frame(
  diagnosis = rep("Meningioma", 5),
  sex = rep("Female", 5),
  loc = rep("Supratentorial", 5),
  ki = unique(bc2$ki),
  gtv = rep(mean(bc2$gtv), 5),
  stereo = rep("SRT", 5)
)


survplots <- survfit(fit.all, newdata = modaldata2)

plot(survplots, col=2:6)
legend("bottomleft", levels(bc$diagnosis), col=2:6, lty = 1)
```
 
 
 
# Practice on Dolphin Data
 
 

```{r}

df <- cetaceans::cetaceans

df |> 
  ggplot(aes(age, sex)) +
  ggridges::geom_density_ridges(aes(fill=sex)) +
  theme_grey(base_size = 18)

df |> 
  ggplot(aes(age, as.factor(event))) +
  ggridges::geom_density_ridges() +
  theme_grey(base_size = 18)


```



```{r}


res.cox <- coxph(
  Surv(age, event) ~ species 
  #+ sex 
  + transfers 
  + strata(born_in_captivity)
  , data = df
  )


res.cox <- coxph(
  Surv(age, event) ~ species 
  #+ sex 
  + transfers 
  + born_in_captivity
  , data = df
  )
res.cox |> 
  broom::tidy()


df_pred <- df |> 
  count(species) |> 
  mutate(transfers = mean(df$transfers),
         born_in_captivity = mean(df$born_in_captivity))


fit.new <- survfit(res.cox, newdata = df_pred)

plot(fit.new, col = 2:21)
legend("topright", levels(df_pred$species), col=2:21, legend = df_pred$species, lty = 1)

fit.new |> 
  broom::tidy() |> 
  select(time, starts_with('estimate')) |> 
  pivot_longer(cols = starts_with('estimate')) |> 
  left_join(
    df_pred |> 
      mutate(rn = row_number(), 
             name = paste0('estimate.', rn)) |> 
      select(name, species)
  ) |> 
  ggplot(aes(time, value)) +
  geom_line(aes(color = species), alpha = 0.5) +
  theme(legend.position = 'bottom')
  

```







