---
title: "Untitled"
format: html
editor: visual
---



```{r}
library(tidyverse)

library(censored)


```


```{r}

bc <- ISLR2::BrainCancer |> 
  as_tibble()

```

Data:
- **status**: 0 alive, 1 dead
- **gtv**: gross turmor volume (cubic cm)
- **sex**: male or female 
- **diagnosis**: meningioma LG glioma, HG glioma, or other 
- **loc**: tumor location infratentorial or supratentorial 
- **ki**: Karnofsky index 
- **stereo**: stereotactic method; either stereotactic radiosurgery or fractionated stereotactic radiotherapy, abbreviated as SRS and SRT, respectively


53 of 88 patients were still alive at the end. 

```{r}
bc |> 
  count(status) |> 
  mutate(total = sum(n),
         perc = n / total)

```



```{r}
table(bc$sex, bc$status)
```

Sex did not show statistical significance in difference with p=0.2

```{r}
library(survival)

fit.surv <- survfit(Surv(time, status) ~ 1, data=bc)
plot(fit.surv, xlab='months', ylab='Est. prob of survival')

fit.sex <- survfit(Surv(time, status) ~ sex, data = bc)
plot(fit.sex, col = c(2, 4))
legend("bottomleft", levels(bc$sex), col = c(2,4), lty = 1)

logrank.test <- survdiff(Surv(time, status) ~ sex, data=bc)
logrank.test

fit.cox <- coxph(Surv(time, status) ~ sex, data = bc)
summary(fit.cox)

```

all predictors

```{r}
fit.all <- coxph(
  Surv(time, status) ~ 
    sex 
  + diagnosis
  + loc
  + ki 
  + gtv
  + stereo
  , data = bc
    )


fit.all


modaldata <- data.frame(
  diagnosis = levels(bc$diagnosis),
  sex = rep("Female", 4),
  loc = rep("Supratentorial", 4),
  ki = rep(mean(bc$ki), 4),
  gtv = rep(mean(bc$gtv), 4),
  stereo = rep("SRT", 4)
)


survplots <- survfit(fit.all, newdata = modaldata)

plot(survplots, col=2:5)
legend("bottomleft", levels(bc$diagnosis), col=2:5, lty = 1)

```

## Publication data

```{r}
pub <- ISLR2::Publication |> as_tibble()

fit.posres <- survfit(
  Surv(time, status) ~ posres, data = pub
)
fit.posres$surv

fit.pub <- coxph(Surv(time, status) ~ posres, data = pub)
fit.pub

logrank.test.pub <- survdiff(Surv(time, status) ~ posres, data = pub)
logrank.test.pub


```

```{r}
fit.pub2 <- coxph(Surv(time, status) ~ . - mech, data=pub)
fit.pub2

fit.pub3 <- coxph(Surv(time, status) ~ posres  + clinend + budget + impact, data = pub)
fit.pub3
```
 
 
```{r}
ex_4 <- tibble(
  observation = c(26.5, 37.2, 57.3, 90.8, 20.2, 89.8),
  censor_ind = c(1, 1, 1, 0, 0, 0),
  covariate_x = c(0.1, 11, -0.3, 2.8, 1.8, 0.4)
)

coxph(Surv(observation, censor_ind) ~ covariate_x, data = ex_4) 

y <- survfit(Surv(observation, censor_ind) ~ 1, data = ex_4)
plot(y)


```
Applied
exercise 10
 
```{r}

res <- survfit(Surv(time, status) ~ 1# sex + diagnosis + loc + ki + gtv + stereo
               , data=bc)
plot(res)
```
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 

```{r}

df <- cetaceans::cetaceans

df |> 
  ggplot(aes(age, sex)) +
  ggridges::geom_density_ridges(aes(fill=sex)) +
  theme_grey(base_size = 18)

df |> 
  ggplot(aes(age, as.factor(event))) +
  ggridges::geom_density_ridges() +
  theme_grey(base_size = 18)


df |> 
  count(age, event) |> 
  ggplot(aes(age, n)) +
  geom_col() +
  facet_wrap(~event, ncol = 1)

```



```{r}


res.cox <- coxph(
  Surv(age, event) ~ species 
  #+ sex 
  + transfers 
  + born_in_captivity
  , data = df
  )

test.ph <- cox.zph(res.cox)
test.ph



```







